Watch Out for the Bully! Job Interference Study on Dragonfly Network

### 2016-04-03 hybrid_placement.tex
cut off tables in section 4,5,6. reduce num of pages to 12

We propose a hybrid job placement policy to assign each job with preferable allocation. 
According to the hybrid placement policy, 
the less communication-intensive jobs get contiguous allocation to partially avoid network sharing with the ``bully", 
while the communication-intensive jobs get random allocation in order to share network resource with each other. 
When Workload~\Rmnum{1} is running on the dragonfly network with hybrid placement policy, 
AMG gets contiguous allocation, MultiGrid and CrystalRouter get random allocations. 
The experiment results show that hybrid placement performs comparably as random placement for improving network performance and mitigates the degradation of AMG caused by the ``bullying".

\subsection{Network Performance Analysis}

Hybrid placement is also coupled with three routing policies, minimal, adaptive and progressive adaptive, denoted respectively as HM, HA and HPA. As presented in section~\ref{sec: workload-1 network analysis} and~\ref{sec: workload-2 network analysis}, random placement outperforms contiguous placement in terms of network performance, the analysis in this section only focuses on the comparison between hybrid and random placement policies.

\begin{table}[ht]
\begin{center}
\caption{Average time spent on communication by all MPI ranks when Workload I is running on dragonfly network under hybrid placement and random placement policies.} 
\label{tab: hyb-placement-wkld-commtime}
\begin{tabular}{l c c c c c c }
\toprule % Top horizontal line
\toprule
&\multicolumn{6}{c}{Placement and Routing Configurations} \\
\cmidrule(l){2-7}
          & HM & HA & HPA & RM & RA & RPA \\ % Column names row
\midrule % In-table horizontal line
Time(ms)  &273 &255 &255 &255 &265 &264  \\ % Content row 1
%\midrule
%Workload II &1747 &1991 &1991 &1791 &2367 &1965 \\
\midrule % In-table horizontal line
\bottomrule % Bottom horizontal line
\end{tabular}
\end{center}
\end{table}

Due to page limit, we only present the average communication time spent by Workload~\Rmnum{1} for the evaluation of the network performance.
\footnote{The analysis about network traffic and saturated time are omitted in this section due to page limit.}
As shown in Table~\ref{tab: hyb-placement-wkld-commtime}, 
hybrid placement coupled with (progressive) adaptive routing (HA, HPA) can guarantee the same performance as random placement coupled with minimal routing (RM), thus the average workload communication time stays unchanged. 
When hybrid placement coupled with minimal routing (HM) is in use, 
the network performance declines  as indicated by the increasing average communication time. 
The consecutive groups assigned to AMG results in the diminishing of network resource shared by MultiGrid and CrystalRouter, 
causing unbalanced utilization over the network.

\subsection{Individual Application Analysis}

>>>>

We propose a hybrid job placement policy to assign each job with preferable allocation. 
According to the hybrid placement policy, 
the less communication-intensive jobs get contiguous allocation to partially avoid network sharing with the ``bully", 
while the communication-intensive jobs get random allocation in order to share network resource with each other. 
When Workload~\Rmnum{1} is running on the dragonfly network with hybrid placement policy, 
AMG gets contiguous allocation, MultiGrid and CrystalRouter get random allocations. 
The experiment results show that hybrid placement performs comparably as random placement for improving network performance and mitigates the degradation of AMG caused by the ``bullying".
\footnote{The analysis about network traffic and saturated time are omitted in this section due to page limit.}

%\subsection{Network Performance Analysis}
%
%Hybrid placement is also coupled with three routing policies, minimal, adaptive and progressive adaptive, denoted respectively as HM, HA and HPA. As presented in section~\ref{sec: workload-1 network analysis} and~\ref{sec: workload-2 network analysis}, random placement outperforms contiguous placement in terms of network performance, the analysis in this section only focuses on the comparison between hybrid and random placement policies. 
%
%\begin{table}[ht]
%\begin{center}
%\caption{Average time spent on communication by all MPI ranks when Workload I is running on dragonfly network under hybrid placement and random placement policies.} 
%\label{tab: hyb-placement-wkld-commtime}
%\begin{tabular}{l c c c c c c }
%\toprule % Top horizontal line
%\toprule
%&\multicolumn{6}{c}{Placement and Routing Configurations} \\
%\cmidrule(l){2-7}
%          & HM & HA & HPA & RM & RA & RPA \\ % Column names row
%\midrule % In-table horizontal line
%Time(ms)  &273 &255 &255 &255 &265 &264  \\ % Content row 1
%%\midrule
%%Workload II &1747 &1991 &1991 &1791 &2367 &1965 \\
%\midrule % In-table horizontal line
%\bottomrule % Bottom horizontal line
%\end{tabular}
%\end{center}
%\end{table}
%
%Due to page limit, we only present the average communication time spent by Workload~\Rmnum{1} for the evaluation of the network performance.
%\footnote{The analysis about network traffic and saturated time are omitted in this section due to page limit.}
%As shown in Table~\ref{tab: hyb-placement-wkld-commtime}, 
%hybrid placement coupled with (progressive) adaptive routing (HA, HPA) can guarantee the same performance as random placement coupled with minimal routing (RM), thus the average workload communication time stays unchanged. 
%When hybrid placement coupled with minimal routing (HM) is in use, 
%the network performance declines  as indicated by the increasing average communication time. 
%The consecutive groups assigned to AMG results in the diminishing of network resource shared by MultiGrid and CrystalRouter, 
%causing unbalanced utilization over the network.

%\subsection{Individual Application Analysis}

### 2016-04-03 intro.tex
cut off tables in section 4,5,6. reduce num of pages to 12

Based on the aforementioned key observations, the optimal job placement policy for the workload running dragonfly system should be specific to the application communication intensity. Thus, we propose a hybrid job placement, which can assign the communication-intensive applications with random allocation and less intensive applications with contiguous allocation. Our experiment results show that hybrid job placement can mitigate the performance degradation of less communication-intensive applications while still guarantee the best performance of communication-intensive applications.
%\emph{The random placement and adaptive routing are preferable to communication-intensive jobs, whereas contiguous placement and minimal routing are in favor of less intensive jobs.}

To the best of our knowledge, using parallel application traces from production systems for the study of job interference on dragonfly networks has not been reported so far. We believe the observations and new placement policy presented in this paper are valuable to a number of communities including HPC computing centers, system software developers, and system administrators. For example, the computing centers should take resource sharing into consideration when choosing proper configurations for building their dragonfly networks. System software developers could design better scheduling algorithms for jobs with distinct communication behavior. System administrators could make more accurate predication about system availability based on job running status under different placement and routing configurations.

>>>>

Based on the aforementioned key observations, the optimal job placement policy for the workload running dragonfly system should be specific to the application communication intensity. Thus, we propose a hybrid job placement, which can assign the communication-intensive applications with random allocation and less intensive applications with contiguous allocation. Our experiment results show that hybrid job placement can mitigate the performance degradation of less communication-intensive applications while still guarantee the best performance of communication-intensive applications.
%\emph{The random placement and adaptive routing are preferable to communication-intensive jobs, whereas contiguous placement and minimal routing are in favor of less intensive jobs.} 
To the best of our knowledge, using real application traces from production systems for the study of job interference on dragonfly networks has not been reported so far. We believe the observations and new placement policy presented in this paper are valuable to the HPC community.
%We believe the observations and new placement policy presented in this paper are valuable to a number of communities including HPC computing centers, system software developers, and system administrators. 
%For example, the computing centers should take resource sharing into consideration when choosing proper configurations for building their dragonfly networks. System software developers could design better scheduling algorithms for jobs with distinct communication behavior. System administrators could make more accurate predication about system availability based on job running status under different placement and routing configurations.

### 2016-04-03 study_workload_1.tex
rewriting abstract and introduction done; tweak format in section 4
cut off tables in section 4,5,6. reduce num of pages to 12

%===========================================================
%from the workload perspective to study the network performance. 
The efficiency of the workload communication is another perspective to evaluate the network performance. 
The average communication time spent by all MPI ranks when Workload~\Rmnum{1} is running under different placement and routing configurations shown in Table~\ref{tab:wkld-commtime}. 
The random placement policy performs comparably with all routing policies (RM, RA and RPA). They outperform all contiguous placement configurations (CM, CA and CPA). The contiguous placement coupled with minimal routing (CM) results in highest communication time. When coupled with (progressive) adaptive routing (CA, CPA), the average communication time can be greatly reduced. When random placement is in use, the workload communication efficiency can be further improved.

\begin{table}[ht]
\begin{center}
\caption{Average communication time by all MPI ranks when Workload I is running on the dragonfly network under six different placement and routing configurations.} 
\label{tab:wkld-commtime}
\begin{tabular}{l c c c c c c }
\toprule % Top horizontal line
\toprule
&\multicolumn{6}{c}{Placement and Routing Configurations} \\ 
\cmidrule(l){2-7}
          & CM & CA & CPA & RM & RA & RPA \\ % Column names row
\midrule % In-table horizontal line
Time(ms)  & 476  & 300  & 301  & 255  & 265  & 264  \\ % Content row 1

\midrule % In-table horizontal line
\bottomrule % Bottom horizontal line
\end{tabular}
\end{center}
\end{table}

The random placement can balance the network traffic load by uniformly distribute the traffic. The (progressive) adaptive routing can avoid the hot-spots by redirecting the traffic away from congested area. The network can reach its optimal performance, when random placement coupled with (progressive) adaptive routing are in use.

@@

AMG is quite an exception. 
The communication time of AMG running with different placement and routing configuration are shown in Figure~\ref{fig:amg-commtime}.
When contiguous placement is in use, AMG behaves similarly as MultiGrid and CrystalRouter over three routing policies. 
It takes comparable amount of communication time when contiguous placement coupled with adaptive and progressive routing. 
When coupled with minimal routing, the communication time increase twofold. 
AMG behaves abnormally when random placement is in use. 
The communication time skyrockets when random placement coupled with (progressive) adaptive routing (RPA, RA) are in use. Minimal routing (RM) can mitigate this performance degradation, but still worse than CA and CPA.

We observe that when Workload~\Rmnum{1} is running with random placement and (progressive) adaptive routing, 
the performance of MultiGrid and CrystalRouter are significant improved, whereas AMG suffers performance degradation.

@@

Since AMG is less communication-intensive, 
the routers serving AMG have small volume of traffic on local and global channels when the contiguous placement is use, 
shown as the dashed line in Figure~\ref{fig:amg-lc-traffic},~\ref{fig:amg-gc-traffic}. 
When random placement is in use, 
more routers are involved in serving AMG, 
and the traffic skyrockets on the local and global channels, 
shown as the solid lines in Figure~\ref{fig:amg-lc-traffic},~\ref{fig:amg-gc-traffic}. 
The increased traffic comes from the communication-intensive MultiGrid and CrystalRouter.

AMG behaves abnormally when random placement is in use 
because the routers serving AMG have to take the traffic from MultiGrid and CrystalRouter. 
The extra traffic from the other communication-intensive applications 
takes advantage the network resources belonging to AMG and slows down AMG's communication. 
We refers this phenomenon as AMG being ``bullied" by MultiGrid and CrystalRouter.

>>>>

%===========================================================
%from the workload perspective to study the network performance. 
%The efficiency of the workload communication is another perspective to evaluate the network performance. 
%The average communication time spent by all MPI ranks when Workload~\Rmnum{1} is running under different placement and routing configurations shown in Table~\ref{tab:wkld-commtime}. 
%The random placement policy performs comparably with all routing policies (RM, RA and RPA). They outperform all contiguous placement configurations (CM, CA and CPA). The contiguous placement coupled with minimal routing (CM) results in highest communication time. When coupled with (progressive) adaptive routing (CA, CPA), the average communication time can be greatly reduced. When random placement is in use, the workload communication efficiency can be further improved. 
%
%
%\begin{table}[ht]
%\begin{center}
%\caption{Average communication time by all MPI ranks when Workload I is running on the dragonfly network under six different placement and routing configurations.} 
%\label{tab:wkld-commtime}
%\begin{tabular}{l c c c c c c }
%\toprule % Top horizontal line
%\toprule
%&\multicolumn{6}{c}{Placement and Routing Configurations} \\ 
%\cmidrule(l){2-7}
%          & CM & CA & CPA & RM & RA & RPA \\ % Column names row
%\midrule % In-table horizontal line
%Time(ms)  & 476  & 300  & 301  & 255  & 265  & 264  \\ % Content row 1
%
%\midrule % In-table horizontal line
%\bottomrule % Bottom horizontal line
%\end{tabular}
%\end{center}
%\end{table}
%
%
%The random placement can balance the network traffic load by uniformly distribute the traffic. The (progressive) adaptive routing can avoid the hot-spots by redirecting the traffic away from congested area. The network can reach its optimal performance, when random placement coupled with (progressive) adaptive routing are in use. 
%

@@

AMG is quite an exception. 
The communication time of AMG running with different placement and routing configuration are shown in Figure~\ref{fig:amg-commtime}.
When contiguous placement is in use, AMG behaves similarly as MultiGrid and CrystalRouter over three routing policies. 
It takes comparable amount of communication time when contiguous placement coupled with adaptive and progressive routing. 
When coupled with minimal routing, the communication time almost doubles. 
The communication time skyrockets when random placement coupled with (progressive) adaptive routing (RPA, RA) are in use. Minimal routing (RM) can mitigate this performance degradation, but still worse than CA and CPA. We observe that when Workload~\Rmnum{1} is running with random placement and (progressive) adaptive routing, 
the performance of MultiGrid and CrystalRouter are significant improved, whereas AMG suffers performance degradation.

@@

Since AMG is less communication-intensive, 
the routers serving AMG have small volume of traffic on local and global channels when the contiguous placement is use, 
shown as the dashed line in Figure~\ref{fig:amg-lc-traffic},~\ref{fig:amg-gc-traffic}. 
When random placement is in use, 
more routers are involved in serving AMG, 
and the traffic skyrockets on the local and global channels, 
shown as the solid lines in Figure~\ref{fig:amg-lc-traffic},~\ref{fig:amg-gc-traffic}. 
The increased traffic comes from the communication-intensive MultiGrid and CrystalRouter. 
AMG behaves abnormally when random placement is in use 
because the routers serving AMG have to take the traffic from MultiGrid and CrystalRouter. 
The extra traffic from the other communication-intensive applications 
takes advantage the network resources belonging to AMG and slows down AMG's communication. 
We refers this phenomenon as AMG being ``bullied" by MultiGrid and CrystalRouter.

### 2016-04-03 study_workload_2.tex
cut off tables in section 4,5,6. reduce num of pages to 12

The average communication time spent by all MPI ranks when Workload~\Rmnum{2} is running under different placement and routing configurations shown in Table~\ref{tab:syn-wkld-commtime}. 
We can observe the similar pattern as in Table~\ref{tab:wkld-commtime}
The contiguous placement coupled with minimal routing results in highest communication time. 
The random placement coupled with minimal routing is the most efficient. 
Progressive adaptive routing is better than adaptive routing when random placement is in use, 
and they have comparable results when coupled contiguous placement. 
Same as Workload~\Rmnum{1}, the network reaches the optimal performance when Workload~\Rmnum{2} is running with random placement and adaptive routing.

\begin{table}[ht]
\begin{center}
\caption{Average time spent on communication by all MPI ranks when Workload II is running on dragonfly network under six different placement and routing configurations.} 
\label{tab:syn-wkld-commtime}
%\centering % Centers the table on the page, comment out to left-justify
\begin{tabular}{l c c c c c c }
\toprule % Top horizontal line
\toprule
&\multicolumn{6}{c}{Placement and Routing Configuration} \\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
\cmidrule(l){2-7}
          & CM & CA & CPA & RM & RA & RPA \\ % Column names row
\midrule % In-table horizontal line
Time(ms)  & 3360 & 2204 & 2335 & 1791 & 2367 & 1965 \\ % Content row 1

\midrule % In-table horizontal line
\bottomrule % Bottom horizontal line
\end{tabular}
\end{center}
\end{table}

>>>>

%+================================================================
%The average communication time spent by all MPI ranks when Workload~\Rmnum{2} is running under different placement and routing configurations shown in Table~\ref{tab:syn-wkld-commtime}. 
%We can observe the similar pattern as in Table~\ref{tab:wkld-commtime}
%The contiguous placement coupled with minimal routing results in highest communication time. 
%The random placement coupled with minimal routing is the most efficient. 
%Progressive adaptive routing is better than adaptive routing when random placement is in use, 
%and they have comparable results when coupled contiguous placement. 
%Same as Workload~\Rmnum{1}, the network reaches the optimal performance when Workload~\Rmnum{2} is running with random placement and adaptive routing. 
%
%\begin{table}[ht]
%\begin{center}
%\caption{Average time spent on communication by all MPI ranks when Workload II is running on dragonfly network under six different placement and routing configurations.} 
%\label{tab:syn-wkld-commtime}
%%\centering % Centers the table on the page, comment out to left-justify
%\begin{tabular}{l c c c c c c }
%\toprule % Top horizontal line
%\toprule
%&\multicolumn{6}{c}{Placement and Routing Configuration} \\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
%\cmidrule(l){2-7}
%          & CM & CA & CPA & RM & RA & RPA \\ % Column names row
%\midrule % In-table horizontal line
%Time(ms)  & 3360 & 2204 & 2335 & 1791 & 2367 & 1965 \\ % Content row 1
%
%\midrule % In-table horizontal line
%\bottomrule % Bottom horizontal line
%\end{tabular}
%\end{center}
%\end{table}
